Embedded Scalars, Preferred Readings \& Prosody: {A}n Experimental Revisit

### 2015-02-06 text/paper/revision_01/paper.tex
K2 reworked section 3 traditionalism

Given this plurality of hypothesized readings, the question arises:
how salient are these theoretically conceivable readings
compared to each other (if they are detectable by empirical means at
all)?  Addressing this question empirically is relevant because it
lies at the heart of a current debate about the nature of quantity
inferences. %  and the demarkation between grammar (i.e. syntax/semantics)
% and domain-general reasoning procedures (i.e. pragmatics).
The controversy is usually seen as one between two main competing
theoretical positions \citep[see][for a detailed comparison of the
two]{ChemlaSingh2014:Remarks-on-the-}.  Simplifying things for now,
these two types of theories make clearly different predictions about
how salient the local readings, in particular, should be.  The first
position which we refer to as \mymark{traditionalism} (similar to, but
not identical to Chemla \& Singh’s ``Gricean approach'') holds that
local readings should not be salient at all. The second one which we
call \mymark{grammaticalism} (Chemla \& Singh’s ``grammatical
approach'') holds that local readings are unproblematically available
and maybe even at least as salient as the literal ones (see Section
\ref{sec:theories-predictions} for a more thorough exposition of the
theories and their predictions).

That matters actually are not as simple as just described is reflected
by lively discussions about which conclusions are permitted by previous
experimental research \citep[e.g.][]{GeurtsPouscoulous2009:Embedded-Implic,CliftonDube2010:Embedded-Implic,ChemlaSpector2010:Experimental-Ev,BenzGotzner2014:Embedded-implic}. The results of these studies
were discussed controversially because of several specific reasons. Firstly,
and most importantly, previous studies presented target sentences visually.
However, traditionalists often acknowledge the availability of local readings for
prosodically marked utterances, albeit not as the result of
a standard scalar implicature but of a different pragmatic process triggered by
prosodic markedness \citep[e.g.][]{Horn2006:The-Border-Wars,Geurts2009:Scalar-Implicat,ChemlaSpector2010:Experimental-Ev,Geurts2010:Quantity-Implic,Tielvan-Tiel2012:Embedded-Scalar,GeurtsTielvan-Tiel2013:Embedded-Scalar}. Under this view, evidence for the availability of local readings \citep[see][]{CliftonDube2010:Embedded-Implic,ChemlaSpector2010:Experimental-Ev} may be
an effect of “silent prosody”  \citep[see e.g.][]{Bader98,Fodor98}. Secondly, as \citet{Tielvan-Tiel2012:Embedded-Scalar} and \citet{GeurtsTielvan-Tiel2013:Embedded-Scalar} argue,
the alleged evidence for the availability of local readings of \as-sentences, might well be
confounded with “typicality effects”  (see Section~\ref{sec:previous-studies} for discussion).
Finally, previous studies have only accumulated limited evidence pertaining to the relative
salience of candidate readings of sentences like (\ref{bsp:AE}) and (\ref{bsp:GE}) (see
Section~\ref{sec:previous-studies} for discussion).

More generally, two prerequisites have to be met in order to make use
of experimental data to decide between theoretical positions regarding
the computation of quantity inferences. Firstly, both traditionalism
and grammaticalism have to be augmented with additional assumptions to
derive testable predictions about the relative salience of candidate
readings. For example, hypothezised effects of prosodic marking have
to be taken into account before we can test traditionalism empirically
using sentences (\ref{bsp:AE}) and (\ref{bsp:GE}).  We will address
this extant position as the \mymark{prosodic markedness hypothesis}
and argue that, if prosody is hypothesized to play a central role, it
is vital to present target sentences auditorily or to control for
effects of “silent prosody” that subjects may apply when reading a
sentence. Secondly, the relative salience of candidate readings has to
be operationalized properly, i.e., we have to devise methods to
measure relative salience. For example, typicality effects and logical
dependencies between readings have to be controlled for when
evaluating speaker judgments in terms of the salience of candidate
readings.  Both of these prerequisites are related to formulating what
\citet{ChemlaSingh2014:Remarks-on-the-} call ``auxiliary assumptions''
that are necessary to enable theoretical accounts to make predictions
about processing. Our emphasis here is not on processing accounts, but
on predictions of behavioral data --- what is sometimes, perhaps
misleadingly, called a computational-level perspective. But even so,
traditionalism and grammaticalism have to summon auxiliary assumptions
which connect the core theories to the dependent variables of a
behavioral experiment. Without proper auxiliary assumptions it is hard
(or impossible) to test traditionalism or grammaticalism empirically.

This paper presents results form an experimental study designed to
deal with the mentioned problems. The focus is on testing certain
formulations of traditionalism and grammaticalsim that are explicit
enough to derive testable predictions about the relative salience of
candidate readings. For practical reasons, we must choose auxiliary
assumptions that actually deliver \emph{ex ante} predictions, even
though this may not exhaust the space of variations on either core
theory that have been floated in the recent literature (see
Section~\ref{sec:theories-predictions} for discussion). In order to
address the prosodic markedness hypothesis, we presented sentence
materials auditorily and explicitly manipulated contrastive stress on
embedded scalar items. In order operationalize the relative salience
of candidate readings, we employed an \mymark{incremental verification
  task} in which parts of a to-be-evaluated picture were hidden and
only revealed piece-wise at the participants’ request until they felt
able to give a binary truth-value judgement
\citep[see][]{Conroy2008}. To analyze the obtained data in terms of
relative salience of candidate readings, we furthermore employed a
statistical model of how the salience of readings guided participants
judgments in the incremental verification task. We validated this
method by including control conditions testing ambiguous sentences
with a well-known preference on available readings. We argue that this
method provides categorical behavioral data that is both clearly
indicative of the reading preferences among participants and largely
independent of potential pictorial confounds, such as the mentioned
typicality effects.

The data obtained by this method is \emph{prima facie} problematic for
both traditionalism and grammaticalism, at least if augmented by the
auxiliary assumptions which we adopted here. However, we want to
stress from the outset that, while the experimental results allow us
to derive restrictions on which auxiliary assumptions are conceivable
for traditionalism and grammaticalism, definite conclusions about the
general validity of the two core theories will be limited
\citep[cf.][]{ChemlaSingh2014:Remarks-on-the-}. Bearing this in mind,
we can summarize our main findings crudely as follows. Contrary to the
predictions of common versions of traditionalism, we observed many
answers indicative of local readings of \es-sentences, irrespective of
prosodic markedness. But also prominent versions of grammaticalism are
incongruent with the observed data.  Backed up by the
\mymark{strongest meaning hypothesis}
\citep{DalrympleKanazawa1998:Reciprocal-Expr} as a disambiguation
criterion \citep[as suggested by][]{ChierchiaFox2008:The-Grammatical},
grammaticalism predicts local readings to be much more salient than we
found evidence for in our data. More strongly even, the preference
order among putative readings suggested by our data cannot even in
principle be explained by comparisons of logical strength among
putative readings. This finding is independently relevant in the light
of the fact that the strongest meaning hypothesis is frequently
summoned as a disambiguation criterion also in other domains, such as
the interpretation of reciprocals
\citep{DalrympleKanazawa1998:Reciprocal-Expr} or vague predicates
\citep{CobrerosEgre2012:Tolerant-Classi}. In sum, our data shed doubt
on both traditionalism and grammaticalism under prominent auxiliary
assumptions. We conclude the paper by some thoughts on how either
position could be amended in the light of our data.

The remainder of the paper is structured as
follows. Section~\ref{sec:get-know-your} elaborates on the three kinds
of relevant readings for our target
sentences. Section~\ref{sec:theories-predictions} works out the
predictions of traditionalism and grammaticalism (plus prominent auxiliary assumptions) 
about availability and relative salience of readings for these
sentences. Section~\ref{sec:previous-studies} reviews related
experimental studies to the extent necessary to motivate our own
approach. Section~\ref{sec:exp} eventually contributes our
experimental design, analyses and results.

@@

There are two main theoretical positions which make different
predictions about the readings of \as- and \es-sentences
\citep[see][for
overview]{Horn2006:The-Border-Wars,Geurts2010:Quantity-Implic,Sauerland2012:The-Computation,ChemlaSingh2014:Remarks-on-the-}. We
will refer to these as \mymark{traditionalism} and
\mymark{grammaticalism} and treat each one in turn.  As already
mentioned in the introduction, there is some leeway in assessing the
predictions for each approach depending on which auxiliary assumptions
to adopt. Therefore, we will distinguish different varieties of either
``core theory.'' While there are certainly other varieties
conceivable, they are, in our perception, the most prominent ones,
but, more importantly, also the most practical ones in that they give
us precise and testable \emph{ex ante} predictions about expected
behavioral data (as opposed to \emph{ex post} explanations after
inspecting the data). The predictions of the considered core theories
plus their auxiliary assumptions are summarized in
Table~\ref{tab:predictions} on page~\pageref{tab:predictions}. Below,
in Section~\ref{sec:critical-reflection}), further conceivable
varieties of traditionalism and grammaticalism will be discussed
\emph{after} inspecting the relevant data.

@@

Traditionalism gets its name from its conservative stance towards
Grice's original theory of conversational implicatures
\citep{Grice1975:Logic-and-Conve}. Many authors have defended
traditionalist positions in this sense. Of the more recent literature,
we would consider as traditionalist, among others, contributions by
\citet{Spector2006:Scalar-Implicat},
\citet{Sauerland2004:Scalar-Implicat},
\citet{Russell2006:Against-Grammat},
\citet{vanRooijSchulz:ExhaustiveInterpretation},
\citet{Geurts2010:Quantity-Implic},
\citet{Franke2011:Quantity-Implic}, or \citet{GoodmanStuhlmuller2013:Knowledge-and-I}.

According to Grice, conversational implicatures, of which quantity
implicatures are a special case, are to be thought of as
rationalizations of speaker behavior. Central in this reasoning is the
assumption that the speaker's behavior is efficient (if not optimal)
and goal-oriented. Usually, the assumed goal of conversation is the
cooperative exchange of relevant information from speaker to
hearer.\footnote{Our conception of ``traditionalism'' is less
  stringent that \citeauthor{ChemlaSingh2014:Remarks-on-the-}'s
  ``Gricean approach.'' We allow our traditionalism to violate the
  constraint that all considered alternatives must be stronger than
  the one to be interpreted, so as to allow, for instance, global
  readings of \es-sentences where the alternative
  (\ref{bsp:GE-Alternative}) does not asymmetrically entail the
  to-be-interpreted (\ref{bsp:GE}) (see more below). We would even
  subsume under traditionalism abductive reasoning to the best
  explanation of what lexical meaning a speaker had intended to convey
  on an occasion of use in an embedded context, if such an account
  could be formulated consistently and in such a manner that it makes
  testable predictions.}

Consequently, the \mymark{Gricean recipe}
\citep[see][]{Geurts2010:Quantity-Implic} for deriving a simple scalar
inference like that in (\ref{bsp:Plain-SI-Implicature}) from an
utterance of (\ref{bsp:Plain-SI-Target}) is as follows: if the issue
whether Hans solved only some or all of the problems is relevant, then
a cooperative and knowledgable speaker would utter
(\ref{bsp:Plain-SI-Alternative}) if in a position to do so; hence, one
of the most natural explanations of why such a speaker has not uttered
(\ref{bsp:Plain-SI-Alternative}), but only (\ref{bsp:Plain-SI-Target})
is that she is uncertain of whether (\ref{bsp:Plain-SI-Alternative})
is true; but on the assumption that she is knowledgeable (competent,
opinionated, informed \dots) it follows that
(\ref{bsp:Plain-SI-Implicature}) should in fact be true.\fn{We are
  glossing here somewhat swiftly over the more nuanced details of the
  derivation of implicatures targeting the speaker's epistemic state
  \citep[e.g.][]{Gazdar1979:Pragmatics:-Imp,Soames1982:How-Presupposit}. Section~\ref{sec:discussion-1}
  will come back to this in more detail.}

@@

As for \as- and \es-sentences, the Gricean recipe derives global
readings in a straightforward way. Consequently, traditionalism
predicts that both literal and global readings are available:  literal
readings, because these are the result of semantic interpretation
and form the starting point of pragmatic reasoning; global readings
because these may be arrived at by the Gricean recipe.

Traditionalism's predictions about relative salience of literal and
global readings depend on whether certain auxiliary assumptions
necessary to derive global readings by the Gricean recipe are
plausibly met. These auxiliary assumptions include relevance of the
extra information provided by the global reading, mutual awareness
that the stronger alternative has been a speaker option, the speaker's
competence about the issue, etc. Normally, traditionalist accounts
would assume that these extra assumptions are met. In that case,
traditionalism would predict that global readings should be preferred
over literal readings. On the other hand, if these extra assumptions
are implausible, traditionalism would predict that literal readings
are preferred over global ones \citep[see][for experimental evidence
indicating that manipulation of speaker knowledgeability indeed
affects the rate of quantity implicatures drawn by the listener
for unembedded occurrences of \emph{some}]{GoodmanStuhlmuller2013:Knowledge-and-I}.

Consequently, there are at least two varieties of traditionalism that
yield slightly different predictions, depending on which extra
assumptions we make about the relevant context of interpretation. If
the context (e.g., our experimental design) makes the extra
assumptions of the Gricean recipe needed to derive non-epistemic
global readings sufficiently salient, then we get the prediction that
global readings should be preferred over literal ones; otherwise the
prediction is that literal readings are preferred over global ones. We
call the former position \mymark{the strong variety of traditionalism}
and the latter the \mymark{the weak variety of traditionalism}.\fn{We
  mention for clarity that although weak traditionalism does not
  predict literal readings to be more salient than non-epistemic
  global readings, it might still predict an \emph{epistemically weak
    implicature}, similar to the global reading, that the speaker is
  uncertain whether the stronger alternative is true. (See also the
  discussion in Section~\ref{sec:discussion-1})} It is of course an
empirical task to determine when a context would favor either variety
of traditionalism. For our purposes, however, it suffices to note that
the core theory is plausibly compatible with either preference
relation: literal over global or the other way around (see also
Table\ref{tab:predictions}).

Does traditionalism also predict local readings to be available? The
answer is slightly different for \as- and the \es-sentences. For
\es-sentences, traditionalism does not predict that a local reading is
available, at least not as a quantity implicature by conjoining the
literal meaning with a suitable set of negated alternatives
\citep[see][]{GeurtsPouscoulous2009:Embedded-Implic,ChemlaSpector2010:Experimental-Ev}. The
literal and the local reading of \es-sentences are logically
independent. But if $X$ and $Y$ are logically independent
propositions, then there is no proposition $Z$ such that $X$ would be
equivalent to $Y \wedge \neg Z$.

On the other hand, as for \as-sentences, there is a traditionalist way
of deriving the local reading, namely by assuming that not only
(\ref{bsp:AE-Alternative}) is an alternative to (\ref{bsp:AE}), but
also the sentence in (\ref{bsp:AE-Alternatives-Extended}) \citep[e.g.][]{Sauerland2004:Scalar-Implicat}.

\begin{exer}{bsp:AE}
  \ex \mymark{All} of the students solved \mymark{some} of the problems.
\end{exer}

\begin{exer}{bsp:AE-Alternative}
  \ex \label{bsp:AE-Alternatives-Standard}\mymark{All} of the students solved \mymark{all} of the problems.
\end{exer}

@@

\noindent Clearly, if we conjoin a literal reading of (\ref{bsp:AE})
with the negation of (\ref{bsp:AE-Alternatives-Extended}) we obtain
exactly the local reading. (Notice that the negation of
(\ref{bsp:AE-Alternatives-Extended}) entails the negation of
(\ref{bsp:AE-Alternative}), so it makes no difference (not) to add it
in this case.) Consequently, traditionalism does predict that the
local reading of \as-sentences is available if
(\ref{bsp:AE-Alternatives-Extended}) is an available alternative. We
therefore introduce another distinction: \mymark{restricted
  traditionalism} assumes that (\ref{bsp:AE-Alternatives-Extended}) is
not available and therefore predicts that the local reading for
\as-sentences is not available; in contrast, \mymark{unrestricted
  traditionalism} assumes that (\ref{bsp:AE-Alternatives-Extended}) is
available and so is the local reading.\fn{The term traditionalism may
  be misleading here because traditionalists usually assume that
  scalar implicatures should only be computed when a sentence is
  uttered in a context in which a logically stronger alternative is
  relevant \cite{ChemlaSingh2014:Remarks-on-the-}. However, the
  alternative in (\ref{bsp:AE-Alternatives-Extended}) is not stronger
  than (\ref{bsp:AE}). For completeness sake, we still decided to
  subsume “unrestricted traditionalism” under traditionalism because,
  once alternatives like (\ref{bsp:AE-Alternatives-Extended}) are
  acknowledged, the local reading of \as-sentences can be derived in a
  straightforward traditionalist manner \citep[see also][who call this
  type of theory ``unrestricted
  globalism'']{ChemlaSpector2010:Experimental-Ev}.}

As far as preferences over readings of \as-sentences are concerned,
even unrestricted traditionalism would not predict that the local
reading is preferred over the global reading. This is because the
derivation of the local reading via
(\ref{bsp:AE-Alternatives-Extended}) hinges on yet another auxiliary
assumption, namely the availability of the alternative in
(\ref{bsp:AE-Alternatives-Extended}), which seems, by comparing
numbers of replacements, much less obvious and therefore presumably is
less readily available than the alternative in
(\ref{bsp:AE-Alternative}). Consequently, we derive the following
predictions about preferences for unrestricted traditionalism: weak
unrestricted traditionalism prefers the literal over the global over
the local reading, while strong unrestricted traditionalism prefers
the global over the local over the literal reading (see also
Table~\ref{tab:predictions}).

@@

\begin{table}[t]
  \centering
  \begin{tabular}{lcc}
    & \multicolumn{2}{c}{preference \& availability}
    \\ \cmidrule(r){2-3}
    theoretical position
    & \as
    & \es
    \\ \midrule
    traditionalism
    \\
    \ \ \ \ weak restricted 
    & \lit > \glb 
    & \lit > \glb
    \\
    \ \ \ \ weak unrestricted
    & \lit > \glb > \loc 
    & \lit > \glb
    \\
    \ \ \ \ strong restricted
    & \glb > \lit 
    & \glb > \lit
    \\
    \ \ \ \ strong unrestricted
    & \glb > \loc > \lit 
    & \glb > \lit
    \\
    % conventionalism
    % \\
    % \ \ defaultism
    % & \loc >  \lit 
    % & \loc >  \lit
    % \\
    % \ \ literalism
    % & \lit >  \loc 
    % & \lit >  \loc
    % \\
    grammaticalism
    \\
    \ \ \ \ weak
    & \loc > \glb > \lit 
    & \glb > \lit, \loc
    \\
    \ \ \ \ strong
    & \glb, \loc > \lit 
    & \glb, \loc , \lit
    \\
  \end{tabular}
  \caption{Predictions of the relevant theoretical positions. Listed
    are the predicted preference relations. If a reading is not
    listed, it is predicted to be unavailable.}
  \label{tab:predictions}
\end{table}

>>>>

Given this plurality of hypothesized readings, the question arises:
how salient are these theoretically conceivable readings compared to
each other (if they are detectable by empirical means at all)?
Addressing this question empirically is relevant because it lies at
the heart of a current debate about the nature of quantity inferences.
The controversy is usually seen as one between two main competing
theoretical positions, which we call \mymark{traditionalism} and
\mymark{grammaticalism} and take to be general ideas behind large
classes of approach to quantity implicatures \citep[c.f.][for related
discussion with a focus on deriving processing
theories]{ChemlaSingh2014:Remarks-on-the-}. Unfortunately, it is
rather difficult to empirically test either ``core theory'' (
terminology from \citeauthor{ChemlaSingh2014:Remarks-on-the-})
directly because additional ``auxiliary assumptions'' are needed in
both cases to derive falsifiable \emph{ex ante} predictions about
to-be-expected observations in a concrete behavioral experiment, even
irrespective of processing accounts. We therefore try to carefully set
up plausible instantiations of either core theory with particular
auxiliary assumptions that give testable predictions at the outset in
Section~\ref{sec:theories-predictions}, then evaluate the obtained
predictions against our data, and finally broaden our perspective to
reflect in Section~\ref{sec:critical-reflection} on the impact of our
data on the more general tenability of the core theories.

But even with crisp theoretical predictions at hand, matters are still
very complex, as reflected by lively discussions about which
conclusions are permitted by previous experimental research
\citep[e.g.][]{GeurtsPouscoulous2009:Embedded-Implic,CliftonDube2010:Embedded-Implic,ChemlaSpector2010:Experimental-Ev,BenzGotzner2014:Embedded-implic}. The
results of these studies were discussed controversially because of
several specific reasons. Firstly, and most importantly, previous
studies presented target sentences visually.  However, traditionalists
often acknowledge the availability of local readings for prosodically
marked utterances, albeit not as the result of a standard scalar
implicature but of a different pragmatic process triggered by prosodic
markedness
\citep[e.g.][]{Horn2006:The-Border-Wars,Geurts2009:Scalar-Implicat,ChemlaSpector2010:Experimental-Ev,Geurts2010:Quantity-Implic,Tielvan-Tiel2012:Embedded-Scalar,GeurtsTielvan-Tiel2013:Embedded-Scalar}. Under
this view, evidence for the availability of local readings
\citep[see][]{CliftonDube2010:Embedded-Implic,ChemlaSpector2010:Experimental-Ev}
may be an effect of “silent prosody” \citep[see
e.g.][]{Bader98,Fodor98}. We will therefore address this extant
position as the \mymark{prosodic markedness hypothesis} and argue
that, if prosody is hypothesized to play a central role, it is vital
to present target sentences auditorily or to control for effects of
“silent prosody” that subjects may apply when reading a
sentence. Secondly, as \citet{Tielvan-Tiel2012:Embedded-Scalar} and
\citet{GeurtsTielvan-Tiel2013:Embedded-Scalar} argue, the alleged
evidence for the availability of local readings of \as-sentences,
might well be confounded with, in particular, “typicality effects”
(see Section~\ref{sec:previous-studies} for discussion).  Finally,
previous studies have only accumulated limited evidence pertaining to
the relative salience of candidate readings of sentences like
(\ref{bsp:AE}) and (\ref{bsp:GE}) (see
Section~\ref{sec:previous-studies} for discussion).

% More generally, two prerequisites have to be met in order to make use
% of experimental data to decide between theoretical positions regarding
% the computation of quantity inferences. Firstly, both traditionalism
% and grammaticalism have to be augmented with additional assumptions to
% derive testable predictions about the relative salience of candidate
% readings. For example, hypothezised effects of prosodic marking have
% to be taken into account before we can test traditionalism empirically
% using sentences (\ref{bsp:AE}) and (\ref{bsp:GE}).  We will address
% this extant position as the \mymark{prosodic markedness hypothesis}
% and argue that, if prosody is hypothesized to play a central role, it
% is vital to present target sentences auditorily or to control for
% effects of “silent prosody” that subjects may apply when reading a
% sentence. Secondly, the relative salience of candidate readings has to
% be operationalized properly, i.e., we have to devise methods to
% measure relative salience. For example, typicality effects and logical
% dependencies between readings have to be controlled for when
% evaluating speaker judgments in terms of the salience of candidate
% readings.  Both of these prerequisites are related to formulating what
% \citet{ChemlaSingh2014:Remarks-on-the-} call ``auxiliary assumptions''
% that are necessary to enable theoretical accounts to make predictions
% about processing. Our emphasis here is not on processing accounts, but
% on predictions of behavioral data --- what is sometimes, perhaps
% misleadingly, called a computational-level perspective. But even so,
% traditionalism and grammaticalism have to summon auxiliary assumptions
% which connect the core theories to the dependent variables of a
% behavioral experiment. Without proper auxiliary assumptions it is hard
% (or impossible) to test traditionalism or grammaticalism empirically.

To deal with these methodological problems, this paper presents
results form an experimental study which presented sentence materials
auditorily and explicitly manipulated contrastive stress on embedded
scalar items. We employed an \mymark{incremental verification task} in
which parts of a to-be-evaluated picture were hidden and only revealed
piece-wise at the participants’ request until they felt able to give a
binary truth-value judgement \citep[see][]{Conroy2008}. We use both a
classical regression analysis and a Bayesian statistical model to
analyze the data from this novel task, and validate this method by
including control conditions testing ambiguous sentences with a
well-known preference on available readings. We argue that this method
provides categorical behavioral data that is both clearly indicative
of the reading preferences among participants and largely independent
of potential pictorial confounds, such as the mentioned typicality
effects.

% The data obtained by this method is \emph{prima facie} problematic for
% both traditionalism and grammaticalism, at least if augmented by the
% auxiliary assumptions which we adopted here. However, we want to
% stress from the outset that, while the experimental results allow us
% to derive restrictions on which auxiliary assumptions are conceivable
% for traditionalism and grammaticalism, definite conclusions about the
% general validity of the two core theories will be limited
% \citep[cf.][]{ChemlaSingh2014:Remarks-on-the-}. Bearing this in mind,
% we can summarize our main findings crudely as follows. Contrary to the
% predictions of common versions of traditionalism, we observed many
% answers indicative of local readings of \es-sentences, irrespective of
% prosodic markedness. But also prominent versions of grammaticalism are
% incongruent with the observed data.  Backed up by the
% \mymark{strongest meaning hypothesis}
% \citep{DalrympleKanazawa1998:Reciprocal-Expr} as a disambiguation
% criterion \citep[as suggested by][]{ChierchiaFox2008:The-Grammatical},
% grammaticalism predicts local readings to be much more salient than we
% found evidence for in our data. More strongly even, the preference
% order among putative readings suggested by our data cannot even in
% principle be explained by comparisons of logical strength among
% putative readings. This finding is independently relevant in the light
% of the fact that the strongest meaning hypothesis is frequently
% summoned as a disambiguation criterion also in other domains, such as
% the interpretation of reciprocals
% \citep{DalrympleKanazawa1998:Reciprocal-Expr} or vague predicates
% \citep{CobrerosEgre2012:Tolerant-Classi}. In sum, our data shed doubt
% on both traditionalism and grammaticalism under prominent auxiliary
% assumptions. We conclude the paper by some thoughts on how either
% position could be amended in the light of our data.

The paper is structured as follows. Section~\ref{sec:get-know-your}
elaborates on the three kinds of relevant readings for our target
sentences. Section~\ref{sec:theories-predictions} works out plausible
\emph{ex ante} predictions of the core theories traditionalism and
grammaticalism (for some practically relevant auxiliary assumptions)
about availability and relative salience of readings for our target
sentences. Section~\ref{sec:previous-studies} reviews related
experimental studies to the extent necessary to motivate our own
approach. Section~\ref{sec:exp} eventually contributes our
experimental design, analyses and results. Finally,
Section~\ref{sec:critical-reflection} discusses our method and
interpretation of the results in a broader theoretical context.

@@

There are two main theoretical positions which appear to make
different predictions about the readings of \as- and \es-sentences
\citep[see][for
overview]{Horn2006:The-Border-Wars,Geurts2010:Quantity-Implic,Sauerland2012:The-Computation,ChemlaSingh2014:Remarks-on-the-}. We
will refer to these as \mymark{traditionalism} and
\mymark{grammaticalism} and treat each in turn. Both approaches are
addressed here as ``core theories'': two main ideas behind several
approaches to quantity implicatures. To determine concrete predictions
from these, additional ``auxiliary assumptions'' are necessary. Here,
we spell a set of such auxiliary assumptions for each core theory. Our
motivation for the particular choices we make is partly guided by
practical concerns, so as to have a concrete set of predictions for
each core theory to formulate expectations and compare our data to. We
therefore introduce some prominent or otherwise relevant variants of
either core theory, the predictions of which are summarized in
Table\ref{tab:predictions} on page~\pageref{tab:predictions}. How
other potential instantiations of either core theory fare in light of
our data will be discussed in Section~\ref{sec:critical-reflection}.

@@

We use ``traditionalism'' as a vague umbrella term for approaches that
build on \citeauthor{Grice1975:Logic-and-Conve}'s
(\citeyear{Grice1975:Logic-and-Conve}) original notion that
conversational implicatures, of which quantity implicatures are a
special case, are to be thought of as rationalizations of speaker
behavior. Central in this reasoning is the assumption that the
speaker's behavior is efficient (if not optimal) and
goal-oriented. Usually, the assumed goal of conversation is the
cooperative exchange of relevant information from speaker to
hearer.\footnote{Of the more recent literature, we would consider as
  traditionalist, among many others, contributions by
  \citet{Spector2006:Scalar-Implicat},
  \citet{Sauerland2004:Scalar-Implicat},
  \citet{Russell2006:Against-Grammat},
  \citet{vanRooijSchulz:ExhaustiveInterpretation},
  \citet{Geurts2010:Quantity-Implic},
  \citet{Franke2011:Quantity-Implic}, or
  \citet{GoodmanStuhlmuller2013:Knowledge-and-I}.}

The traditionalist view gives rise to what
\citet{Geurts2010:Quantity-Implic} calls the \mymark{Gricean recipe}
for deriving a scalar inference (\ref{bsp:Plain-SI-Implicature}) from
an utterance of (\ref{bsp:Plain-SI-Target}): if the issue whether Hans
solved only some or all of the problems is relevant, then a
cooperative and knowledgable speaker would utter
(\ref{bsp:Plain-SI-Alternative}) if in a position to do so; hence, one
of the most natural explanations of why such a speaker has not uttered
(\ref{bsp:Plain-SI-Alternative}), but only (\ref{bsp:Plain-SI-Target})
is that she is uncertain of whether (\ref{bsp:Plain-SI-Alternative})
is true; but on the assumption that she is knowledgeable (competent,
opinionated, informed \dots) it follows that
(\ref{bsp:Plain-SI-Implicature}) should in fact be true.\fn{We are
  glossing here somewhat swiftly over the more nuanced details of the
  derivation of implicatures targeting the speaker's epistemic state
  \citep[e.g.][]{Gazdar1979:Pragmatics:-Imp,Soames1982:How-Presupposit}. More
  on this below.}

@@

Traditionalism's predictions about availability and relative salience
of readings of \as- and \es-sentences depend on further auxiliary
assumptions. To have handy labels, we distinguish between strong and
weak variants, and between restricted and unrestricted
variants.

Strong and weak traditionalism differ with respect to whether the
assumptions needed to derive non-epistemic implicatures are prominent
or not in the given context of utterance (or some experiment or
other). These assumptions include relevance of the information in
question, mutual awareness that the stronger alternative has been a
speaker option, the speaker's competence about the issue,
etc. \emph{Strong traditionalism} assumes that the additional
assumptions needed to derive a non-epistemic implicatures are
sufficiently salient to prefer the non-epistemic implicature over the
literal reading; \emph{weak traditionalism} assumes the reverse
preference order. In other words, in the context of a given
experiment, it is compatible with a traditionalist position to expect
literal or implicature readings to be more prominent. Or, put yet
differently, a traditionalist could be uncertain as to whether a given
experiment is more likely to trigger more implicature readings or
literal readings. Without more information about the experiment or
context of interpretation, traditionalism is compatible with either
salience ordering.

Orthogonal to this, we distinguish restricted and unrestricted
variants of traditionalism. For our purposes, restricted
traditionalism considers only alternative expressions that are
stronger than the to-be-interpreted utterance. Unrestricted
traditionalism is more relaxed and also allows alternatives that are
derived from replacing \emph{some} with \emph{all} and vice versa, if
necessary multiple times, even if the resulting alternative is not
logically stronger. Restricted traditionalism considers only
\as-sentences to have a relevant alternative, namely
(\ref{bsp:AE-Alternative}). It then predicts the availability of a
global reading for \as-sentences on top of the literal. Restricted
traditionalism predicts only literal readings of \es-sentences,
because the alternative (\ref{bsp:GE-Alternative}) does not entail the
literal reading of (\ref{bsp:GE}). Every implicature reading available
to restricted traditionalism is also available to unrestricted
traditionalism, but the latter's predictions are even richer. For
\as-sentences, the potential alternative
(\ref{bsp:AE-Alternatives-Extended}) also makes local readings
available \citep[e.g.][]{Sauerland2004:Scalar-Implicat}: conjoining a
literal reading of (\ref{bsp:AE}) with the negation of
(\ref{bsp:AE-Alternatives-Extended}) gives a local reading.

@@

\noindent For \es-sentences, we derive a global reading via the
negation of alternative (\ref{bsp:GE-Alternative}).

In sum, a traditionalist core theory is compatible with a range of
observations about the availability and relative salience of
implicature readings. These are summarized in
Table\ref{tab:predictions}. Most importantly for the subsequent
discussion, traditionalism does not predict local readings for
\es-sentences and can favor implicature readings over literal ones or
vice versa.

@@

\begin{table}[t]
  \centering
  \begin{tabular}{lcc}
    & \multicolumn{2}{c}{preference \& availability}
    \\ \cmidrule(r){2-3}
    theoretical position
    & \as
    & \es
    \\ \midrule
    traditionalism
    \\
    \ \ \ \ weak restricted 
    & \lit > \glb 
    & \lit 
    \\
    \ \ \ \ weak unrestricted
    & \lit > \glb , \loc 
    & \lit > \glb
    \\
    \ \ \ \ strong restricted
    & \glb > \lit 
    & \lit 
    \\
    \ \ \ \ strong unrestricted
    & \glb , \loc > \lit 
    & \glb > \lit
    \\
    % conventionalism
    % \\
    % \ \ defaultism
    % & \loc >  \lit 
    % & \loc >  \lit
    % \\
    % \ \ literalism
    % & \lit >  \loc 
    % & \lit >  \loc
    % \\
    grammaticalism
    \\
    \ \ \ \ weak
    & \loc > \glb > \lit 
    & \glb > \lit, \loc
    \\
    \ \ \ \ strong
    & \glb, \loc > \lit 
    & \glb, \loc , \lit
    \\
  \end{tabular}
  \caption{Predictions of the relevant theoretical positions. Listed
    are the predicted preference relations. If a reading is not
    listed, it is predicted to be unavailable. Comma-separated
    readings are not ranked, i.e., not predicted to differ as opposed
    to being predicted to be equally ranked.}
  \label{tab:predictions}
\end{table}
